---
params: 
  figures_title: "Light Rail and Tram Publication Figures"
  figures_author: "Rose McNally"
  publication_fin_year: ~
  minimal_tidy_dataset: ~
title: "`r params$figures_title`"
author: "`r params$figures_author`"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

previous_fin_year <- tramlr::get_previous_fin_year(publication_fin_year)
```

## Front page

```{r, front-page, include=FALSE}

# Get passenger journeys change, this year and difference figures for England
# Get passenger journeys change and this year figures for London and England outside London

passenger_journeys_index <- grep("LRT0101", names(params$minimal_tidy_dataset))
passenger_journeys <- params$minimal_tidy_dataset[[passenger_journeys_index]]
passenger_journeys <- utils::tail(dplyr::select(passenger_journeys, London, England, `England outside of London`), 2)

England_passenger_journeys <- passenger_journeys$England[[2]]
England_last_passenger_journeys <- passenger_journeys$England[[1]]
England_diff_passenger_journeys <- England_passenger_journeys - England_last_passenger_journeys
England_change_passenger_journeys <- England_diff_passenger_journeys * 100 / England_last_passenger_journeys

London_passenger_journeys <- passenger_journeys$London[[2]]
London_last_passenger_journeys <- passenger_journeys$London[[1]]
London_diff_passenger_journeys <- London_passenger_journeys - London_last_passenger_journeys
London_change_passenger_journeys <- London_diff_passenger_journeys * 100 / London_last_passenger_journeys

EOL_passenger_journeys <- passenger_journeys$`England outside of London`[[2]]
EOL_last_passenger_journeys <- passenger_journeys$`England outside of London`[[1]]
EOL_diff_passenger_journeys <- EOL_passenger_journeys - EOL_last_passenger_journeys
EOL_change_passenger_journeys <- EOL_diff_passenger_journeys * 100 / EOL_last_passenger_journeys


```

The were `r tramlr::round_up(England_passenger_journeys, 1)` million passenger journeys in England in `r publication_fin_year`. The difference from `r previous_fin_year` is `r tramlr::round_up(England_diff_passenger_journeys, 1)` million, a percentage change of `r tramlr::round_up(England_change_passenger_journeys, 1)`%.

In London, there were `r tramlr::round_up(London_passenger_journeys, 1)` million passenger journeys, a percentage change of `r tramlr::round_up(London_change_passenger_journeys, 1)`%.

In England outside of London, there were `r tramlr::round_up(EOL_passenger_journeys, 1)` million passenger journeys, a percentage change of `r tramlr::round_up(EOL_change_passenger_journeys, 1)`%.

## Factsheet

```{r, factsheet, include=FALSE}

# Passenger journeys per head in England this year

journeys_per_head_index <- grep("LRT0109", names(params$minimal_tidy_dataset))
journeys_per_head <- params$minimal_tidy_dataset[[journeys_per_head_index]]
England_journeys_per_head <- utils::tail(journeys_per_head$England, 1)

# Concessionary journeys in England this year

concession_journeys_index <- grep("LRT0102", names(params$minimal_tidy_dataset))
concession_journeys <- params$minimal_tidy_dataset[[concession_journeys_index]]
England_concession_journeys <- utils::tail(concession_journeys$England, 1)

# Percent of all journeys that were concessionary in England

passenger_concession_percent <- England_concession_journeys * 100 / England_passenger_journeys

# Gross revenue this year in England

gross_revenue_index <- grep("LRT0301a", names(params$minimal_tidy_dataset))
gross_revenue <- params$minimal_tidy_dataset[[gross_revenue_index]]
England_gross_revenue <- utils::tail(gross_revenue$England, 1)

# Concessionary revenue this year in England

concession_revenue_index <- grep("LRT0302a", names(params$minimal_tidy_dataset))
concession_revenue <- params$minimal_tidy_dataset[[concession_revenue_index]]
England_concession_revenue <- utils::tail(concession_revenue$England, 1)

# Average revenue per passenger journey in England this year

England_av_revenue_per_journey <- England_gross_revenue / England_passenger_journeys

# Average revenue per concessionary journey in England this year

England_av_revenue_per_cons <- England_concession_revenue / England_concession_journeys

# Revenue percentage change this year in England

England_last_gross_revenue <- utils::tail(gross_revenue$England, 2)[[1]]
England_diff_gross_revenue <- England_gross_revenue - England_last_gross_revenue
England_change_gross_revenue <- England_diff_gross_revenue * 100 / England_last_gross_revenue

# Percentage of passenger revenue from DLR this year in England

England_DLR_gross_revenue <- utils::tail(gross_revenue$`Docklands Light Railway`, 1)
England_DLR_percent_gross_revenue <- England_DLR_gross_revenue * 100 / England_gross_revenue

```

There were `r tramlr::round_up(England_passenger_journeys, 1)` million passenger journeys and `r tramlr::round_up(England_journeys_per_head, 1)` passenger journeys per head in England in `r publication_fin_year`. Passenger journeys changed by `r tramlr::round_up(England_change_passenger_journeys, 1)`%.

There were `r tramlr::round_up(England_concession_journeys, 1)` million concessionary journeys, and they made up `r tramlr::round_up(passenger_concession_percent, 1)`% of all journeys.

The gross revenue from passenger journeys was `r paste("\U00A3", tramlr::round_up(England_gross_revenue, 1), sep = "")` million, of which `r paste("\U00A3", tramlr::round_up(England_concession_revenue, 1), sep = "")` million was concessionary revenue. The average revenue per passenger journey was `r paste("\U00A3", tramlr::round_up(England_av_revenue_per_journey, 2), sep = "")`, and the average concessionary revenue per concessionary journey was `r paste("\U00A3", tramlr::round_up(England_av_revenue_per_cons, 2), sep = "")`. Revenue changed by `r tramlr::round_up(England_change_gross_revenue, 1)`% on the previous year. `r tramlr::round_up(England_DLR_percent_gross_revenue)`% of passenger revenue came from Docklands Light Railway.

## Summary figures

```{r, summary-figures, echo=FALSE}

# Passenger journey table for last two years across Enlgand operators

passenger_journeys <- params$minimal_tidy_dataset[[passenger_journeys_index]]
passenger_journeys <- utils::tail(passenger_journeys, 2)
passenger_journeys_table <- dplyr::select(passenger_journeys, `Financial year`:`Blackpool Tramway`)

knitr::kable(passenger_journeys_table,
               caption = "Passenger Journeys compared with previous year (millions)",
               align = "lrrrrrrrr",
               col.names = names(passenger_journeys_table))

# Vehicle miles table for last two years across Enlgand operators

vehicle_miles_index <- grep("LRT0106", names(params$minimal_tidy_dataset))
vehicle_miles <- params$minimal_tidy_dataset[[vehicle_miles_index]]
vehicle_miles <- utils::tail(vehicle_miles, 2)
vehicle_miles_table <- dplyr::select(vehicle_miles, `Financial year`:`Blackpool Tramway`)

knitr::kable(vehicle_miles_table,
               caption = "Vehicle miles compared with previous year (millions)",
               align = "lrrrrrrrr",
               col.names = names(vehicle_miles_table))

# Average journey length in miles in England this year

av_journey_length_index <- grep("LRT0107a", names(params$minimal_tidy_dataset))
av_journey_length <- params$minimal_tidy_dataset[[av_journey_length_index]]

England_av_journey_length <- utils::tail(av_journey_length$England, 1)
England_av_journey_length <- measurements::conv_unit(England_av_journey_length, "km", "mi") 

# London

London_av_journey_length <- utils::tail(av_journey_length$London, 1)
London_av_journey_length <- measurements::conv_unit(London_av_journey_length, "km", "mi")

# England outside of London

EOL_av_journey_length <- utils::tail(av_journey_length$`England outside of London`, 1)
EOL_av_journey_length <- measurements::conv_unit(EOL_av_journey_length, "km", "mi")

```

The average light rail and tram journey in England was `r tramlr::round_up(England_av_journey_length, 1)` miles. In London it was `r tramlr::round_up(London_av_journey_length, 1)` miles, and in England outside of London it was `r tramlr::round_up(EOL_av_journey_length, 1)` miles.

## Light rail and tram safety

See ORR data.

## Infrastructure

```{r, infrastructure, echo = FALSE}

# Number of stops table for last two years across Enlgand operators

no_of_stops_index <- grep("LRT0201", names(params$minimal_tidy_dataset))
no_of_stops <- params$minimal_tidy_dataset[[no_of_stops_index]]
no_of_stops <- utils::tail(no_of_stops, 2)
no_of_stops_table <- dplyr::select(no_of_stops, `Financial year`:`Blackpool Tramway`)

knitr::kable(no_of_stops_table,
               caption = "Number of stops compared with previous year",
               align = "lrrrrrrrr",
               col.names = names(no_of_stops_table))

# Number of trams table for last two years across Enlgand operators

no_of_trams_index <- grep("LRT0202", names(params$minimal_tidy_dataset))
no_of_trams <- params$minimal_tidy_dataset[[no_of_trams_index]]
no_of_trams <- utils::tail(no_of_trams, 2)
no_of_trams_table <- dplyr::select(no_of_trams, `Financial year`:`Blackpool Tramway`)

knitr::kable(no_of_trams_table,
               caption = "Number of carriages compared with previous year",
               align = "lrrrrrrrr",
               col.names = names(no_of_trams_table))

# Route miles table for last two years across Enlgand operators

route_miles_index <- grep("LRT0204", names(params$minimal_tidy_dataset))
route_miles <- params$minimal_tidy_dataset[[route_miles_index]]
route_miles <- utils::tail(route_miles, 2)
route_miles_table <- dplyr::select(route_miles, `Financial year`:`Blackpool Tramway`)

knitr::kable(route_miles_table,
               caption = "Route miles compared with previous year",
               align = "lrrrrrrrr",
               col.names = names(route_miles_table))

# Route miles in Engand this year

England_route_miles <- utils::tail(route_miles$England, 1)

# Percentage change

England_last_route_miles <- route_miles$England[[1]]
England_diff_route_miles <- England_route_miles - England_last_route_miles
England_change_route_miles <- England_diff_route_miles * 100 / England_last_route_miles

# Number of stops in England this year

England_no_of_stops <- utils::tail(no_of_stops$England, 1)

# Percentage change

England_last_no_of_stops <- no_of_stops$England[[1]]
England_diff_no_of_stops <- England_no_of_stops - England_last_no_of_stops
England_change_no_of_stops <- England_diff_no_of_stops * 100 / England_last_no_of_stops

# Number of trams in England this year

England_no_of_trams <- utils::tail(no_of_trams$England, 1)

# Percentage change

England_last_no_of_trams <- no_of_trams$England[[1]]
England_diff_no_of_trams <- England_no_of_trams - England_last_no_of_trams
England_change_no_of_trams <- England_diff_no_of_trams * 100 / England_last_no_of_trams

# Percentage change in passenger journeys by network in England this year

TWM_diff_passenger_journeys <- passenger_journeys_table$`Tyne and Wear Metro`[[2]] - passenger_journeys_table$`Tyne and Wear Metro`[[1]]
TWM_change_passenger_journeys <- TWM_diff_passenger_journeys * 100 / passenger_journeys_table$`Tyne and Wear Metro`[[1]] 

Sheffield_diff_passenger_journeys <- passenger_journeys_table$`Sheffield Supertram`[[2]] - passenger_journeys_table$`Sheffield Supertram`[[1]]
Sheffield_change_passenger_journeys <- Sheffield_diff_passenger_journeys * 100 / passenger_journeys_table$`Sheffield Supertram`[[1]] 

NET_diff_passenger_journeys <- passenger_journeys_table$`Nottingham Express Transit`[[2]] - passenger_journeys_table$`Nottingham Express Transit`[[1]]
NET_change_passenger_journeys <- NET_diff_passenger_journeys * 100 / passenger_journeys_table$`Nottingham Express Transit`[[1]] 

DLR_diff_passenger_journeys <- passenger_journeys_table$`Docklands Light Railway`[[2]] - passenger_journeys_table$`Docklands Light Railway`[[1]]
DLR_change_passenger_journeys <- DLR_diff_passenger_journeys * 100 / passenger_journeys_table$`Docklands Light Railway`[[1]] 

Lon_Tram_diff_passenger_journeys <- passenger_journeys_table$`London Tramlink`[[2]] - passenger_journeys_table$`London Tramlink`[[1]]
Lon_Tram_change_passenger_journeys <- Lon_Tram_diff_passenger_journeys * 100 / passenger_journeys_table$`London Tramlink`[[1]] 

MML_diff_passenger_journeys <- passenger_journeys_table$`Midland Metro`[[2]] - passenger_journeys_table$`Midland Metro`[[1]]
MML_change_passenger_journeys <- MML_diff_passenger_journeys * 100 / passenger_journeys_table$`Midland Metro`[[1]] 

Man_Met_diff_passenger_journeys <- passenger_journeys_table$`Manchester Metrolink`[[2]] - passenger_journeys_table$`Manchester Metrolink`[[1]]
Man_Met_change_passenger_journeys <- Man_Met_diff_passenger_journeys * 100 / passenger_journeys_table$`Manchester Metrolink`[[1]] 

BLP_diff_passenger_journeys <- passenger_journeys_table$`Blackpool Tramway`[[2]] - passenger_journeys_table$`Blackpool Tramway`[[1]]
BLP_change_passenger_journeys <- BLP_diff_passenger_journeys * 100 / passenger_journeys_table$`Blackpool Tramway`[[1]] 


```

The route miles in England was `r tramlr::round_up(England_route_miles, 0)` miles, a change of `r tramlr::round_up(England_change_route_miles, 0)`%.

The number of stations in England was `r tramlr::round_up(England_no_of_stops, 0)`, a change of `r tramlr::round_up(England_change_no_of_stops, 0)`%.

The number of carriages in England was `r tramlr::round_up(England_no_of_trams, 0)`, a change of `r tramlr::round_up(England_change_no_of_trams, 0)`%.

Changes in passenger journeys:

* Tyne and Wear Metro `r tramlr::round_up(TWM_change_passenger_journeys, 1)`%

* Sheffield Supertram `r tramlr::round_up(Sheffield_change_passenger_journeys, 1)`%

* Nottingham Express Transit `r tramlr::round_up(NET_change_passenger_journeys, 1)`%

* Docklands Light Railway `r tramlr::round_up(DLR_change_passenger_journeys, 1)`%

* London Tramlink `r tramlr::round_up(Lon_Tram_change_passenger_journeys, 1)`%

* Midland Metro `r tramlr::round_up(MML_change_passenger_journeys, 1)`%

* Manchester Metrolink `r tramlr::round_up(Man_Met_change_passenger_journeys, 1)`%

* Blackpool Tramway `r tramlr::round_up(BLP_change_passenger_journeys, 1)`%

## Passenger Journeys

```{r, passenger-journeys, include=FALSE}

# Change in passenger journeys from ten years ago

passenger_journeys <- params$minimal_tidy_dataset[[passenger_journeys_index]]
passenger_journeys <- utils::tail(passenger_journeys, ten_year_comparison)

ten_financial_years_ago <- passenger_journeys$`Financial year`[[1]]

ten_years_ago_passenger_journeys <- passenger_journeys$`England`[[1]]
ten_years_ago_diff_passenger_journeys <- England_passenger_journeys - ten_years_ago_passenger_journeys
ten_years_ago_change_passenger_journeys <- ten_years_ago_diff_passenger_journeys * 100 / ten_years_ago_passenger_journeys

# Change in journeys per head from ten years ago

journeys_per_head <- utils::tail(journeys_per_head, ten_year_comparison)
ten_years_ago_journeys_per_head <- journeys_per_head$England[[1]]
ten_years_ago_diff_journeys_per_head <- England_journeys_per_head - ten_years_ago_journeys_per_head
ten_years_ago_change_journeys_per_head <- ten_years_ago_diff_journeys_per_head * 100 / ten_years_ago_journeys_per_head

# Ten years ago journeys per head for London and England outside of London

London_ten_years_ago_journeys_per_head <- journeys_per_head$London[[1]]
EOL_ten_years_ago_journeys_per_head <- journeys_per_head$`England outside of London`[[1]]

# This year journeys per head in London and England outside of London

London_journeys_per_head <- journeys_per_head$London[[ten_year_comparison]]
EOL_journeys_per_head <- journeys_per_head$`England outside of London`[[ten_year_comparison]]


```

Since `r ten_financial_years_ago`, light rail and tram passenger journeys have changed by `r tramlr::round_up(ten_years_ago_change_passenger_journeys)`%.

The average number of journeys per head in England in `r ten_financial_years_ago` was `r tramlr::round_up(ten_years_ago_journeys_per_head, 1)`. It is now `r tramlr::round_up(England_journeys_per_head, 1)`, so it has changed by `r tramlr::round_up(ten_years_ago_change_journeys_per_head, 1)`%.

London was `r tramlr::round_up(London_ten_years_ago_journeys_per_head, 1)` and is now `r tramlr::round_up(London_journeys_per_head, 1)`.

England outside of London was `r tramlr::round_up(EOL_ten_years_ago_journeys_per_head, 1)` and is now `r tramlr::round_up(EOL_journeys_per_head, 1)`.

## Concessionary journeys

```{r, concession-journeys, include=FALSE}

England_last_concession_journeys <- utils::tail(concession_journeys$England, 2)[[1]]
England_diff_concession_journeys <- England_concession_journeys - England_last_concession_journeys
England_change_concession_journeys <- England_diff_concession_journeys * 100 / England_last_concession_journeys

```

In England, there were `r tramlr::round_up(England_concession_journeys, 1)` million concessionary journeys, a change of `r tramlr::round_up(England_change_concession_journeys, 1)`% from the previous year's figure of `r tramlr::round_up(England_last_concession_journeys, 1)` million. 

`r tramlr::round_up(passenger_concession_percent, 1)`% of all light rail and tram passenger journeys were concessionary. Please see the previous publication for last year's figure.

See NTS for other information.

## Vehicle mileage 

```{r, vehicle-mileage, include=FALSE}

# Change in vehicle miles since last year

# England

England_vehicle_miles <- vehicle_miles$England[[2]]
England_last_vehicle_miles <- vehicle_miles$England[[1]]
England_diff_vehicle_miles <- England_vehicle_miles - England_last_vehicle_miles
England_change_vehicle_miles <- England_diff_vehicle_miles * 100 / England_last_vehicle_miles

# London

London_vehicle_miles <- vehicle_miles$London[[2]]
London_last_vehicle_miles <- vehicle_miles$London[[1]]
London_diff_vehicle_miles <- London_vehicle_miles - London_last_vehicle_miles
London_change_vehicle_miles <- London_diff_vehicle_miles * 100 / London_last_vehicle_miles

# England ouside of London

EOL_vehicle_miles <- vehicle_miles$`England outside of London`[[2]]
EOL_last_vehicle_miles <- vehicle_miles$`England outside of London`[[1]]
EOL_diff_vehicle_miles <- EOL_vehicle_miles - EOL_last_vehicle_miles
EOL_change_vehicle_miles <- EOL_diff_vehicle_miles * 100 / EOL_last_vehicle_miles

# Change in vehicle miles over ten years

# England

vehicle_miles <- params$minimal_tidy_dataset[[vehicle_miles_index]]
vehicle_miles <- utils::tail(vehicle_miles, ten_year_comparison)
ten_years_ago_vehicle_miles <- vehicle_miles$England[[1]]
ten_years_ago_diff_vehicle_miles <- England_vehicle_miles - ten_years_ago_vehicle_miles
ten_years_ago_change_vehicle_miles <- ten_years_ago_diff_vehicle_miles * 100 / ten_years_ago_vehicle_miles

# London

London_ten_years_ago_vehicle_miles <- vehicle_miles$London[[1]]
London_ten_years_ago_diff_vehicle_miles <- London_vehicle_miles - London_ten_years_ago_vehicle_miles
London_ten_years_ago_change_vehicle_miles <- London_ten_years_ago_diff_vehicle_miles * 100 / London_ten_years_ago_vehicle_miles

# England outside of London

EOL_ten_years_ago_vehicle_miles <- vehicle_miles$`England outside of London`[[1]]
EOL_ten_years_ago_diff_vehicle_miles <- EOL_vehicle_miles - EOL_ten_years_ago_vehicle_miles
EOL_ten_years_ago_change_vehicle_miles <- EOL_ten_years_ago_diff_vehicle_miles * 100 / EOL_ten_years_ago_vehicle_miles



```

In England, vehicle mileage was `r tramlr::round_up(England_vehicle_miles, 1)` million miles, a change of `r tramlr::round_up(England_change_vehicle_miles, 1)`% from the year before (`r tramlr::round_up(England_last_vehicle_miles, 1)` million miles). Since `r ten_financial_years_ago`, vehicle mileage has changed by `r tramlr::round_up(ten_years_ago_change_vehicle_miles, 1)`%.

In London, vehicle mileage was `r tramlr::round_up(London_vehicle_miles, 1)` million miles, a change of `r tramlr::round_up(London_change_vehicle_miles, 1)`% from the previous figure of `r tramlr::round_up(London_last_vehicle_miles, 1)` million miles. Since `r ten_financial_years_ago`, vehicle mileage has changed by `r tramlr::round_up(London_ten_years_ago_change_vehicle_miles, 1)`%.

In England outside of London, vehicle mileage was `r tramlr::round_up(EOL_vehicle_miles, 1)` million miles, a change of `r tramlr::round_up(EOL_change_vehicle_miles, 1)`% from the previous figure of `r tramlr::round_up(EOL_last_vehicle_miles, 1)` million miles. Since `r ten_financial_years_ago`, vehicle mileage has changed by `r tramlr::round_up(EOL_ten_years_ago_change_vehicle_miles, 1)`%.

## Revenue

```{r, revenue, include=FALSE}

England_last_cons_revenue <- utils::tail(concession_revenue$England, 2)[[1]]
England_diff_cons_revenue <- England_concession_revenue - England_last_cons_revenue
England_change_cons_revenue <- England_diff_cons_revenue * 100 / England_last_cons_revenue

```

Gross revenue has changed by `r tramlr::round_up(England_change_gross_revenue, 1)`% to `r paste("\U00A3", tramlr::round_up(England_gross_revenue, 1), sep = "")` million.

Concessionary revenue has changed by `r tramlr::round_up(England_change_cons_revenue, 1)`% to `r paste("\U00A3", tramlr::round_up(England_concession_revenue, 1), sep = "")` million.
